---
title: "Reducing GFRN Gene Lists for Nanostring"
author: "David Jenkins"
date: "4/10/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: "united"
---

```{r, echo=FALSE}
library(ggplot2)
load("../../results/session_done_tcga.RData")
#bad not used
results$tcga_bad_Bad_pS112 <- NULL

print_corr_df <- function(indf){
  data.frame(Pathway=c(rep(colnames(indf)[1],3),
                     rep(colnames(indf)[2],3),
                     rep(colnames(indf)[3],3),
                     rep(colnames(indf)[4],2),
                     rep(colnames(indf)[5],4),
                     rep(colnames(indf)[6],4),
                     rep(colnames(indf)[7],3)),
           Genes=c(rep(as.numeric(indf[,1:7])[1],3),
                   rep(as.numeric(indf[,1:7])[2],3),
                   rep(as.numeric(indf[,1:7])[3],3),
                   rep(as.numeric(indf[,1:7])[4],2),
                   rep(as.numeric(indf[,1:7])[5],4),
                   rep(as.numeric(indf[,1:7])[6],4),
                   rep(as.numeric(indf[,1:7])[7],3)),
           Protein=c("Akt","PDK1","PDK1_pS241",
                     "Akt","PDK1","PDK1_pS241",
                     "EGFR","EGFR_pY1068","EGFR_pY1173",
                     "HER2","HER2_pY1248",
                     "IRS1","IGF1R","PDK1","PDK1_pS241",
                     "EGFR","EGFR_pY1068","EGFR_pY1173","MEK1",
                     "MEK1","PKC.alpha","PKC.alpha_pS657"),
           ICBP.cor=c(as.numeric(indf[,10:48])[1:8],
                      NA,
                      as.numeric(indf[,10:48])[9:10],
                      NA,
                      as.numeric(indf[,10:48])[11:15],
                      NA,
                      NA,
                      as.numeric(indf[,10:48])[16:18]),
           TCGA.cor=c(as.numeric(indf[,10:48])[19:30],
                      NA,
                      as.numeric(indf[,10:48])[31:39]))
}

compare_two_corrs <- function(optimizedf, comparetodf){
  optdf <- print_corr_df(optimizedf)
  compdf <-  print_corr_df(comparetodf)
  res <- cbind(compdf,data.frame(ICBP.diff=compdf[,4] - optdf[,4],
                                 TCGA.diff=compdf[,5] - optdf[,5]))
}
```

## Optimized Pathways

The GFRN Pathways Optimized in the Genome Medice Paper are as follows:

```{r, results='asis'}
knitr::kable(print_corr_df(results[1096704,]))
```

The seven pathway signatures contain a total of __980 genes__, but because of
overlapping genes in the signatures there are a total of __759 unique genes__ in
this set of signatures.

To port these signatures onto the nanostring, we want to reduce the number of
unique probes that are needed for this set of signatures. For example, here are
the correlations if all signatures had 10 genes (54 unique genes):

```{r, results='asis'}
knitr::kable(compare_two_corrs(results[1096704,], results[70958,]))
```


## Reduce by Summed Correlation {.tabset}

If we sum the correlation to each of the proteins of interest, the optimized
pathway list has a total correlation of 14.8795. We can compare this summed
correlation to the summed correlation values for all smaller pathway list
correlations (there are a total of 1 million possible smaller gene list
combinations based on the ASSIGn results calculated during pathway optimizations:)

```{r, echo=FALSE}
results$sum_all_cor <- rowSums(results[,10:48])

d <- ggplot(results, aes(sum_nooverlap, sum_all_cor))
d + stat_binhex(bins=250) + 
  scale_fill_gradientn(colours = c("darkblue","white","darkred")) + 
  xlab("Number Unique Genes") + 
  ylab("Summed Correlation")
```

### 1 to 50 genes

```{r,  echo=F, results='asis'}
best_cor_per_genelist <-data.frame(numgenes=rep(0,729), maxcor=rep(0,729))
j <- 1
for (i in sort(unique(results$sum_nooverlap))){
  subset <- results[results$sum_nooverlap == i,]
  max(subset$sum_all_cor)
  best_cor_per_genelist[j,] <- c(i,max(subset$sum_all_cor))
  j <- j+1
}

find_optimal_reduced_list <- function(min, max, s=F){
  bestcor <- max(best_cor_per_genelist[which(best_cor_per_genelist$numgenes >=min & best_cor_per_genelist$numgenes <=max),2])
  bestlen <- best_cor_per_genelist[which(best_cor_per_genelist$maxcor == bestcor),1]
  if(s){
    paste("The best pathway list between",min,"and",max, "is", bestlen,"with a summed correlation of",bestcor, sep=" ")
  }else{
    which(results$sum_all_cor == bestcor & results$sum_nooverlap == bestlen)
  }
}

find_optimal_reduced_list(1,50, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(1,50),]))
```

### 51 to 100 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(51,100, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(51,100),]))
```

### 101 - 150 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(101,150, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(101,150),]))
```

### 151 - 200 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(151,200, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(151,200),]))
```

### 201 - 250 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(201,250, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(201,250),]))
```

### 251 - 300 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(251,300, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(251,300),]))
```

### 301 - 400 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(301,400, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(301,400),]))
```

### 401 - 500 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(401,500, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(401,500),]))
```

### 501 - 600 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(501,600, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(501,600),]))
```

### 601 - 700 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list(601,700, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list(601,700),]))
```

## Reduce by Average Correlation {.tabset}

```{r, echo=F}
results$icbp_akt_AVG    <- apply(results[,10:12], 1, mean)
results$icbp_bad_AVG    <- apply(results[,13:15], 1, mean)
results$icbp_egfr_AVG   <- apply(results[,16:17], 1, mean)
results$icbp_her2_AVG   <- apply(results[,18:19], 1, mean)
results$icbp_igf1r_AVG  <- apply(results[,20:22], 1, mean)
results$icbp_krasgv_AVG <- apply(results[,23:24], 1, mean)
results$icbp_raf_AVG    <- apply(results[,25:27], 1, mean)
results$tcga_akt_AVG    <- apply(results[,28:30], 1, mean)
results$tcga_bad_AVG    <- apply(results[,31:33], 1, mean)
results$tcga_egfr_AVG   <- apply(results[,34:36], 1, mean)
results$tcga_her2_AVG   <- apply(results[,37:38], 1, mean)
results$tcga_igf1r_AVG  <- apply(results[,39:41], 1, mean)
results$tcga_krasgv_AVG <- apply(results[,42:45], 1, mean)
results$tcga_raf_AVG    <- apply(results[,46:48], 1, mean)
results$icbp_AVG_SUM <- rowSums(results[,50:56])
results$tcga_AVG_SUM <- rowSums(results[,57:63])
results$all_AVG_SUM  <- rowSums(results[,64:65])

d <- ggplot(results, aes(sum_nooverlap, all_AVG_SUM))
d + stat_binhex(bins=250) + 
  scale_fill_gradientn(colours = c("darkblue","white","darkred")) + 
  xlab("Number Unique Genes") + 
  ylab("Summed Correlation")
```

### 1 to 50 genes

```{r,  echo=F, results='asis'}
best_cor_per_genelist2 <-data.frame(numgenes=rep(0,729), maxcor=rep(0,729))
j <- 1
for (i in sort(unique(results$sum_nooverlap))){
  subset <- results[results$sum_nooverlap == i,]
  max(subset$all_AVG_SUM)
  best_cor_per_genelist2[j,] <- c(i,max(subset$all_AVG_SUM))
  j <- j+1
}

find_optimal_reduced_list2 <- function(min, max, s=F){
  bestcor <- max(best_cor_per_genelist2[which(best_cor_per_genelist2$numgenes >=min & best_cor_per_genelist2$numgenes <=max),2])
  bestlen <- best_cor_per_genelist2[which(best_cor_per_genelist2$maxcor == bestcor),1]
  if(s){
    paste("The best pathway list between",min,"and",max, "is", bestlen,"with a summed correlation of",bestcor, sep=" ")
  }else{
    which(results$all_AVG_SUM == bestcor & results$sum_nooverlap == bestlen)
  }
}

find_optimal_reduced_list2(1,50, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(1,50),]))
```

### 51 to 100 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(51,100, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(51,100),]))
```

### 101 - 150 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(101,150, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(101,150),]))
```

### 151 - 200 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(151,200, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(151,200),]))
```

### 201 - 250 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(201,250, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(201,250),]))
```

### 251 - 300 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(251,300, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(251,300),]))
```

### 301 - 400 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(301,400, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(301,400),]))
```

### 401 - 500 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(401,500, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(401,500),]))
```

### 501 - 600 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(501,600, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(501,600),]))
```

### 601 - 700 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list2(601,700, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list2(601,700),]))
```

## Reduce by Average Correlation, TCGA Only  {.tabset}

```{r, echo=F}
d <- ggplot(results, aes(sum_nooverlap, tcga_AVG_SUM))
d + stat_binhex(bins=250) + 
  scale_fill_gradientn(colours = c("darkblue","white","darkred")) + 
  xlab("Number Unique Genes") + 
  ylab("Summed Correlation")
```


### 1 to 50 genes

```{r,  echo=F, results='asis'}
best_cor_per_genelist3 <-data.frame(numgenes=rep(0,729), maxcor=rep(0,729))
j <- 1
for (i in sort(unique(results$sum_nooverlap))){
  subset <- results[results$sum_nooverlap == i,]
  max(subset$tcga_AVG_SUM)
  best_cor_per_genelist3[j,] <- c(i,max(subset$tcga_AVG_SUM))
  j <- j+1
}

find_optimal_reduced_list3 <- function(min, max, s=F){
  bestcor <- max(best_cor_per_genelist3[which(best_cor_per_genelist3$numgenes >=min & best_cor_per_genelist3$numgenes <=max),2])
  bestlen <- best_cor_per_genelist3[which(best_cor_per_genelist3$maxcor == bestcor),1]
  if(s){
    paste("The best pathway list between",min,"and",max, "is", bestlen,"with a summed correlation of",bestcor, sep=" ")
  }else{
    which(results$tcga_AVG_SUM == bestcor & results$sum_nooverlap == bestlen)
  }
}

find_optimal_reduced_list3(1,50, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(1,50),]))
```

### 51 to 100 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(51,100, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(51,100),]))
```

### 101 - 150 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(101,150, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(101,150),]))
```

### 151 - 200 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(151,200, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(151,200),]))
```

### 201 - 250 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(201,250, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(201,250),]))
```

### 251 - 300 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(251,300, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(251,300),]))
```

### 301 - 400 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(301,400, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(301,400),]))
```

### 401 - 500 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(401,500, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(401,500),]))
```

### 501 - 600 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(501,600, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(501,600),]))
```

### 601 - 700 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list3(601,700, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list3(601,700),]))
```


## Reduce by Average Correlation, ICBP Only {.tabset}

```{r, echo=F}
d <- ggplot(results, aes(sum_nooverlap, icbp_AVG_SUM))
d + stat_binhex(bins=250) + 
  scale_fill_gradientn(colours = c("darkblue","white","darkred")) + 
  xlab("Number Unique Genes") + 
  ylab("Summed Correlation")
```



### 1 to 50 genes

```{r,  echo=F, results='asis'}
best_cor_per_genelist4 <-data.frame(numgenes=rep(0,729), maxcor=rep(0,729))
j <- 1
for (i in sort(unique(results$sum_nooverlap))){
  subset <- results[results$sum_nooverlap == i,]
  max(subset$icbp_AVG_SUM)
  best_cor_per_genelist4[j,] <- c(i,max(subset$icbp_AVG_SUM))
  j <- j+1
}

find_optimal_reduced_list4 <- function(min, max, s=F){
  bestcor <- max(best_cor_per_genelist4[which(best_cor_per_genelist4$numgenes >=min & best_cor_per_genelist4$numgenes <=max),2])
  bestlen <- best_cor_per_genelist4[which(best_cor_per_genelist4$maxcor == bestcor),1]
  if(s){
    paste("The best pathway list between",min,"and",max, "is", bestlen,"with a summed correlation of",bestcor, sep=" ")
  }else{
    which(results$icbp_AVG_SUM == bestcor & results$sum_nooverlap == bestlen)
  }
}

find_optimal_reduced_list4(1,50, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(1,50),]))
```

### 51 to 100 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(51,100, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(51,100),]))
```

### 101 - 150 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(101,150, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(101,150),]))
```

### 151 - 200 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(151,200, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(151,200),]))
```

### 201 - 250 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(201,250, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(201,250),]))
```

### 251 - 300 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(251,300, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(251,300),]))
```

### 301 - 400 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(301,400, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(301,400),]))
```

### 401 - 500 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(401,500, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(401,500),]))
```

### 501 - 600 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(501,600, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(501,600),]))
```

### 601 - 700 genes

```{r,  echo=F, results='asis'}
find_optimal_reduced_list4(601,700, T)

knitr::kable(compare_two_corrs(results[1096704,], results[find_optimal_reduced_list4(601,700),]))
```

